---
title: "SOC_methods"
author: "Adam Reimer"
format:
  html:
    embed-resources: true
editor: visual
---

```{r include = FALSE}
packs <- c("tidyverse", "ggforce", "RcppRoll", "knitr")
lapply(packs, require, character.only = TRUE)

# source functions
function_files <- list.files(path="..\\functions")
lapply(function_files, function(x) source(paste0("..\\functions\\", x)))

vec_lnalpha <- seq(1, 2, length.out = 3)
vec_sigW <- seq(0.25, 0.75, length.out = 3)
vec_phi <- seq(0, .7, length.out = 3)
beta <- 0.0001
df_power <- data.frame(lnalpha = vec_lnalpha,
                       power = c(0.6, 0.7, 0.8)) #how to pick these. High for SOC sims. Should be lower for OPY sims.
input <- 
  expand.grid(lnalpha = vec_lnalpha, sigW = vec_sigW, phi = vec_phi) %>%
  mutate(beta = beta, 
         Smsy = get_Smsy(lnalpha, beta, TRUE, sigma = sigW, phi = phi),
         lb_eggers = Smsy * 0.8, #Egger bounds
         ub_eggers = Smsy * 1.6) %>%
  rowwise() %>%
  mutate(lb_pctMSY = optimise(get_bounds, #'true' OYP bounds
                              1:Smsy, 
                              lnalpha = lnalpha, 
                              beta = beta,
                              pct_MSY = 0.9,
                              correct = TRUE,
                              sigma = sigW,
                              phi = phi)$minimum,
          ub_pctMSY = optimise(get_bounds, 
                              Smsy:(Smsy*5), 
                              lnalpha = lnalpha, 
                              beta = beta,
                              pct_MSY = 0.7,
                              correct = TRUE,
                              sigma = sigW,
                              phi = phi)$minimum) %>%
  ungroup() %>%
  left_join(df_power, by = "lnalpha")

Chinook_age <- c('3' = 0.1, '4' = 0.2, '5' = 0.3, '6' = 0.38, '7' = 0.02)

input_sims = 5000


# Simulation params -------------------------------------------------------
Ricker_plots <- list()
for(i in 1:length(unique(input$lnalpha))){
Ricker_plots[[i]] <- 
  input %>% 
    slice(rep(1:n(), each = 251)) %>% 
    mutate(S = rep(seq(0, 5 * 1 / max(beta), by = 200), times = nrow(input)),
           R = Ricker(lnalpha = get_lnalpha_p(lnalpha, beta, sigW, phi), beta = beta, S = S),
           Rmsy = Ricker(get_lnalpha_p(lnalpha, beta, sigW, phi), beta, Smsy),
           Smax = 1 / beta,
           Rmax = Ricker(lnalpha = get_lnalpha_p(lnalpha, beta, sigW, phi), beta = beta, S = Smax)) %>%
    ggplot(aes(x = S, y = R)) +
      geom_rect(aes(xmin = lb_pctMSY, xmax = ub_pctMSY, ymin = -Inf, ymax = Inf), fill = "grey95") +
      geom_line() +
      geom_abline(intercept = 0, slope = 1, linetype = 2) +
      geom_point(aes(x = Smax, y = Rmax)) +
      geom_segment(aes(x = Smsy, xend = Smsy, y = Smsy, yend = Rmsy), linetype = 3) +
      theme_bw() +
      facet_grid_paginate(paste0("\u03C6: ", phi) ~ paste0("log(\u03B1): ", lnalpha) + paste0("\u03C3: ", sigW), 
                          #scales = "free_y",
                          ncol = 3, nrow = 3, page = i)
}
```

## Methods

A versatile salmon stock-recruit and fishery simulation model is presented. The goal of this document is to present the model, pick a set of stock assessment parameters which are representative of most Alaskan salmon fisheries and demonstrate how the model parameters interact to simulate stock assessment and fishery dynamics. Future documents will use this model to answer relevant question posed by the ADF&G escapement goal policy implementation team (EGPIT).

### Stock Recruit Process

Abundance of salmon is generated by a S-R relationship that describes the number of fish expected to return (the "recruitment") from a given number of spawning fish (the "escapement"). The total expected recruitment $R_{c}$ produced from fish spawning in brood year $c$ follows the Ricker formulation:

$$
R_{c}~=~S_{c}\alpha \textrm{exp}(-\beta S_{c})
$$

where $S_{c}$ is the number of spawners, $\alpha$ (number of recruits per spawner in the absence of density dependence) is a measure of productivity, and $\beta$ is a measure of density dependence.

However, productivity varies among brood years, fluctuating around a central tendency. Time-varying productivity often manifests as serially correlated model residuals, so a lognormal error term with a lag of 1 year was included in the linearized form of the S-R relationship to represent realized recruitment.

$$
\textrm{ln}(R_{c})~=~\textrm{ln}(S_{c})+\textrm{ln}(\alpha)-\beta S_{c} +\phi \nu_{(c-1)}+\epsilon_{W}
$$

where $\phi$ is the lag-1 serial correlation coefficient, $\nu_{c}$ are model residuals defined as

$$
\nu_{c}~=~\textrm{ln}(R_{c})-\textrm{ln}(S_{c})-\textrm{ln}(\alpha)+\beta S_{c}
$$

and the $\epsilon_{W}$ are independently and normally distributed process errors with "white noise" variance $\sigma^2_{W}$.

Age at maturity ($p_{ca}$), which is needed to distribute recruitment across calendar years, is treated as a constant within each brood year. The abundance $N_{ya}$ of age $a$ salmon in calendar year $y$ is the product of the age at maturity scalar and the total return (recruitment) $R_{c}$ for from brood year $c=y−a$:

$$
N_{ya}~=~p_{(y-a)a}R_{(y-a)}
$$

Total run $N_{y}$ during calendar year $y$ is the sum of abundance at age across all ages:

$$
N_{y}~=~\sum_{a} N_{ya}
$$

Instantaneous fishing mortality $F_y$ is a function of the annual harvest rate $U_y$ and used to calculate escapement $S_y$:

$$
S_{y}~=~N_y*F_y
$$

Where $$F_{y}=-\textrm{log}(1-U_y)\textrm{exp}(\epsilon_F)$$ with $\epsilon_{F}~\sim~\textrm{Normal}(0, \sigma^2_{F})$ and

### Harvest process

The annual harvest rate $U_{y}$, was modeled as a function of the observed total run $N_{y}^{obs}$, both bounds of the escapement goal $lb$ and $ub$, and the power of the fishery.

$$
U_{y}~=~\textrm{f}(N_y^{obs},~lb,~ub,~power)
$$

Where $N_y^{obs}=N_y \textrm{exp}(\epsilon_N)$ with $\epsilon_{N}~\sim~\textrm{Normal}(0, \sigma^2_{N})$. To determine the annual harvest rates we might expect under escapement goal management the following algorithm was used:

$$
U_y^{goal} = \textrm{minimum}(\textrm{Uniform}(U_{ub}, U_{lb}),~power)
$$

where

$$
U_{lb} = \textrm{if}(N_Y^{obs} > lb)~\frac{N_y^{obs} - lb}{N_y^{obs}}~\textrm{else}~0
$$

$$
U_{ub} = \textrm{if}(N_Y^{obs} > ub)~\frac{N_y^{obs} - ub}{N_y^{obs}}~\textrm{else}~0
$$

and $power$ represent the maximum harvest rate the fishery is capable of producing.

### Parameter Space for Salmon Populations

#### Ricker Parameters

Based on the assumption that for Alaskan salmon stocks $\textrm{log}(\alpha)$ varies between $1$ and $2$, process error $\sigma$ varies between $0.25$ and $0.75$, and autocorrelation $\phi$ varies between $0$ and $0.7$ we plan to simulate 27 parameter combinations using all combinations of 3 points within each parameters range (minimum, midpoint, maximum) and setting $\beta=0.0001$ for all simulations. The lower bound of the escapement goal was the escapement that predicted a mean return which produced 90% of MSY while the upper bound was was the escapement that predicted a mean return which produced 70% of MSY. Power varied according to the productivity of the stock and was set to $0.6$, $0.7$, or $0.8$ when $\textrm{log}(\alpha)$ was set to $1.0$, $1.5$ and $2.0$ respectively.

```{r echo = FALSE}
#| label: fig-Ricker
#| fig.height: 8
#| fig.width: 8
#| fig.cap: "Stock recruit relationships and escapement goal ranges used to simulate Alaskan salmon population dynamics. The solid black line represents the mean return, dotted dashed line represents MSY, the dot represents R~max~, and the shaded area represents the escapement goal range. The dashed line represents the 1:1 line."
#| layout-nrow: 3
#| fig-subcap: 
#|   - "log(\u03B1) = 1.0"
#|   - "log(\u03B1) = 1.5"
#|   - "log(\u03B1) = 2.0"
Ricker_plots[[1]] + coord_cartesian(xlim = c(0, 20000), ylim = c(0, 20000))
Ricker_plots[[2]] + coord_cartesian(xlim = c(0, 30000), ylim = c(0, 30000))
Ricker_plots[[3]] + coord_cartesian(xlim = c(0, 50000), ylim = c(0, 50000))
```

```{r include = FALSE}
sim_base <- 
  mapply(FUN = simSR_goal,
         lnalpha = input$lnalpha,
         sigW = input$sigW,
         phi = input$phi,
         lb_goal = input$lb_pctMSY,
         ub_goal = input$ub_pctMSY,
         power = input$power,
         MoreArgs = list(beta = beta,
                         age0 = Chinook_age,
                         Sims0 = 1500,
                         sigN = 0,
                         sigF = 0,
                         Hfun = H_goal),
         SIMPLIFY = FALSE) %>%
  do.call("rbind", .)


# * Plot time series --------------------------------------------------------
#writing function here since it has dubious use outside of this single application
plot_ts <- function(dat, lnalpha0){ 
  df <- 
    dat %>%
    select(lnalpha, sigW, phi, sim, lb_goal, ub_goal, S, N) %>%
    filter(lnalpha == lnalpha0) %>%
    pivot_longer(c(S, N), names_to = "stat", values_to = "value")
  
  limit <- quantile(df$value[df$stat == "N"], .98, na.rm = TRUE)
  
      ggplot(df, aes(x = sim, y = value, color = stat)) +
        geom_line(alpha = 0.4) +
        geom_hline(aes(yintercept = lb_goal), linetype = 2) +
        geom_hline(aes(yintercept = ub_goal), linetype = 2) +
        scale_y_continuous(limits = c(0, limit)) +
        scale_x_continuous(name = "Simulation year") +
        facet_grid(paste0("\u03C6: ", phi) ~ paste0("\u03C3: ", sigW)) +
        ggtitle(label = bquote(log(alpha): .(lnalpha0)))
}

```

Examining a time series of simulation results can help us gain an intuitive sense of the model. Because of the color scheme used (pink for total run and teal for escapement) grey areas represent simulations when there was minimal harvestable surplus. Notice that with modest process error ($\sigma=0.25$) and autocorrelation ($\phi=0$) the fishery is predictable and results in consistent harvestable surplus. As process error increases we see more variability in total run including instances of very large runs for which the fishery had insufficient power to harvest within the goal range and instances of smaller runs for which little or no harvestable surplus was available. Autocorrelation also increases variability in total run (resulting in similar patterns) as well as makes large or small run sizes more persistent. Looking across these 3 panels we see similar patterns for all values of log($\alpha$) although larger values result in larger run sizes and more harvestable surplus.

```{r echo = FALSE, warning = FALSE}
#| label: fig-base
#| fig.height: 8
#| fig.width: 8
#| fig.cap: "Simulated total run and escapments for 27 possible stock recruit parameter combinations."
#| layout-nrow: 3
#| fig-subcap: 
#|   - "log(\u03B1) = 1.0"
#|   - "log(\u03B1) = 1.5"
#|   - "log(\u03B1) = 2.0"
plot_ts(sim_base, 1)
plot_ts(sim_base, 1.5) #+ coord_cartesian(xlim = c(500, 600))
plot_ts(sim_base, 2)
```

#### Management Variability

Note that the simulations above were run without estimation error ($\sigma_N=0$) or implementation error ($\sigma_F=0$)... i.e. perfect management. Both are unreasonable assumptions so lets look at more realistic simulation results.

##### Estimation Error

Estimation error refers to variability in our estimation of total run size and can be thought of as the CV of the inseason run size estimate. We have a fairly good feel for this value (i.e. \~0 for a weir, 5-10 % for a counting project with sampling error, \~30-40% for mark recapture, 0-30% for the harvest component). Increasing estimation error to $\sigma_N=0.3$ (shown below for $\textrm{log}(\alpha) = 1.5$) primarily increases the likelihood of over-escapement.

```{r echo = FALSE, warning = FALSE}
#| label: fig-baseN3
#| fig.height: 8
#| fig.width: 8
#| fig.cap: "Simulated total run and escapments for 9 possible stock recruit parameter combinations with estimation error of \u03C3~N~ = 0.30."
sim_baseN3 <- 
  mapply(FUN = simSR_goal,
         lnalpha = input$lnalpha,
         sigW = input$sigW,
         phi = input$phi,
         lb_goal = input$lb_pctMSY,
         ub_goal = input$ub_pctMSY,
         power = input$power,
         MoreArgs = list(beta = beta,
                         age0 = Chinook_age,
                         Sims0 = 1500,
                         sigN = 0.3,
                         sigF = 0,
                         Hfun = H_goal),
         SIMPLIFY = FALSE) %>%
  do.call("rbind", .)

plot_ts(sim_baseN3, 1.5)
```

Simulations at values of $\sigma_N$ between $0$ and $0.3$ confirm our suspicion that when $\sigma_N$ is varied between 0 and 0.3 we see negligible effects on the percentage of under-escapements with the escapement goal algorithm described above. For over-escapements generally less than 10% of the simulated runs were over the upper bound of the escapement goal when $\sigma_N=0.15$ while less than 15% of the simulated runs were over the upper bound of the escapement goal when $\sigma_N=0.3$. This effect was positively correlated with productivity although the pattern is blurred by variability in the SR relationship at high levels of $\sigma$ or $\phi$.

```{r include = FALSE}
#demonstrate behavior of sigN and sigF
sim_baseF15 <- 
  mapply(FUN = simSR_goal,
         lnalpha = input$lnalpha,
         sigW = input$sigW,
         phi = input$phi,
         lb_goal = input$lb_pctMSY,
         ub_goal = input$ub_pctMSY,
         power = input$power,
         MoreArgs = list(beta = beta,
                         age0 = Chinook_age,
                         Sims0 = 1500,
                         sigN = 0,
                         sigF = 0.15,
                         Hfun = H_goal),
         SIMPLIFY = FALSE) %>%
  do.call("rbind", .)

sim_baseF3 <- 
  mapply(FUN = simSR_goal,
         lnalpha = input$lnalpha,
         sigW = input$sigW,
         phi = input$phi,
         lb_goal = input$lb_pctMSY,
         ub_goal = input$ub_pctMSY,
         power = input$power,
         MoreArgs = list(beta = beta,
                         age0 = Chinook_age,
                         Sims0 = 1500,
                         sigN = 0,
                         sigF = 0.3,
                         Hfun = H_goal),
         SIMPLIFY = FALSE) %>%
  do.call("rbind", .)

sim_baseN15 <- 
  mapply(FUN = simSR_goal,
         lnalpha = input$lnalpha,
         sigW = input$sigW,
         phi = input$phi,
         lb_goal = input$lb_pctMSY,
         ub_goal = input$ub_pctMSY,
         power = input$power,
         MoreArgs = list(beta = beta,
                         age0 = Chinook_age,
                         Sims0 = 1500,
                         sigN = 0.15,
                         sigF = 0,
                         Hfun = H_goal),
         SIMPLIFY = FALSE) %>%
  do.call("rbind", .)

sim_baseF15N15 <- 
  mapply(FUN = simSR_goal,
         lnalpha = input$lnalpha,
         sigW = input$sigW,
         phi = input$phi,
         lb_goal = input$lb_pctMSY,
         ub_goal = input$ub_pctMSY,
         power = input$power,
         MoreArgs = list(beta = beta,
                         age0 = Chinook_age,
                         Sims0 = 1500,
                         sigN = 0.15,
                         sigF = 0.15,
                         Hfun = H_goal),
         SIMPLIFY = FALSE) %>%
  do.call("rbind", .)

sim_baseF3N3 <- 
  mapply(FUN = simSR_goal,
         lnalpha = input$lnalpha,
         sigW = input$sigW,
         phi = input$phi,
         lb_goal = input$lb_pctMSY,
         ub_goal = input$ub_pctMSY,
         power = input$power,
         MoreArgs = list(beta = beta,
                         age0 = Chinook_age,
                         Sims0 = 1500,
                         sigN = 0.3,
                         sigF = 0.3,
                         Hfun = H_goal),
         SIMPLIFY = FALSE) %>%
  do.call("rbind", .)

temp <-
  rbind(sim_base, sim_baseF15) %>%
  rbind(sim_baseN15) %>%
  rbind(sim_baseN3) %>%
  rbind(sim_baseF3) %>%
  rbind(sim_baseF15N15) %>%
  rbind(sim_baseF3N3) %>%
  group_by(lnalpha, sigW, phi, sigN, sigF) %>%
  # mutate(eg_range = ifelse(S < lb_goal, "under", 
  #                          ifelse(S > ub_goal, "over", "within")))
  summarize(under = mean(S < lb_goal),
            over = mean(S > ub_goal),
            within = mean(S >= lb_goal & S <= ub_goal)) %>%
  pivot_longer(cols = c(under, over, within),
               names_to = "eg_range",
               values_to = "pct")
```

```{r echo = FALSE, warning = FALSE}
#| label: fig-sigmaN
#| fig.height: 7
#| fig.width: 8
#| fig.cap: "Simulated percentage of under- and over-escapments for 27 possible stock-recruit parameter combinations each with 3 possible values of observation error."
#| layout-nrow: 2
#| fig-subcap: 
#|   - "Under-escapement"
#|   - "Over-escapement"
temp %>%
  filter(sigN %in% c(0, 0.15, 0.3) & sigF == 0) %>%
  filter(eg_range == "under") %>%
  ggplot(aes(x = as.character(sigN), y = pct, group = lnalpha, fill = as.character(lnalpha)))  +
  geom_bar(stat = "identity", position = "dodge") +
  labs(fill = "log(\u03B1)", y = "Percentage") +
  xlab(~ sigma[N]) +
  facet_grid(paste0("\u03C6: ", phi) ~ paste0("\u03C3: ", sigW))

temp %>%
  filter(sigN %in% c(0, 0.15, 0.3) & sigF == 0) %>%
  filter(eg_range == "over") %>%
  ggplot(aes(x = as.character(sigN), y = pct, group = lnalpha, fill = as.character(lnalpha))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(fill = "log(\u03B1)", y = "Percentage") +
  xlab(~ sigma[N]) +
  facet_grid(paste0("\u03C6: ", phi) ~ paste0("\u03C3: ", sigW))
```

##### Implementation error

Implementation error refers to the ability of our fishery managers to achieve the harvest rate they were trying to achieve. There are a lot of factors that go into missed goals (when the fish are there) and distilling that down to one number is going to be difficult. For example, the harvest model described above implies mangers are aiming for any point within the escapement goal range with equal probability but changing this to a different management strategy would effect the results below drastically. That said, increasing implementation error to $\sigma_F=0.3$ (shown below for $\textrm{log}(\alpha) = 1.5$) increases the likelihood of both under- and over-escapement.

```{r echo = FALSE, warning = FALSE}
#| label: fig-baseF3
#| fig.height: 7
#| fig.width: 8
#| fig.cap: "Simulated total run and escapments for 9 possible stock recruit parameter combinations with log(\\u03B1) = 1.5 and implementation error of \u03C3~F~ = 0.30."
plot_ts(sim_baseF3, 1.5)
```

Simulations at values of $\sigma_F$ between $0$ and $0.3$ can quantify increases in the percentage of simulated years with both over and under escapements. In general, less than 5% of the simulated runs missed the lower bound of the escapement goal when $\sigma_F=0.15$ while less than 10% of the simulated runs missed the lower bound of the escapement goal when $\sigma_F=0.3$. This effect was positively correlated with productivity although the pattern is blurred by variability in the SR relationship at high levels of $\sigma$ or $\phi$. Similar patterns occurred for over-escapement.

```{r echo = FALSE, warning = FALSE}
#| label: fig-sigmaF
#| fig.height: 7
#| fig.width: 8
#| fig.cap: "Simulated percentage of under- and over-escapments for 27 possible stock-recruit parameter combinations each with 3 possible values of implentation error."
#| layout-nrow: 2
#| fig-subcap: 
#|   - "Under-escapement"
#|   - "Over-escapement"
temp %>%
  filter(sigF %in% c(0, 0.15, 0.3) & sigN == 0) %>%
  filter(eg_range == "under") %>%
  ggplot(aes(x = as.character(sigF), y = pct, group = lnalpha, fill = as.character(lnalpha))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(fill = "log(\u03B1)", y = "Percentage") +
  xlab(~ sigma[F]) +
  facet_grid(paste0("\u03C6: ", phi) ~ paste0("\u03C3: ", sigW))

temp %>%
  filter(sigF %in% c(0, 0.15, 0.3) & sigN == 0) %>%
  filter(eg_range == "over") %>%
  ggplot(aes(x = as.character(sigF), y = pct, group = lnalpha, fill = as.character(lnalpha)))  +
  geom_bar(stat = "identity", position = "dodge") +
    labs(fill = "log(\u03B1)", y = "Percentage") +
  xlab(~ sigma[F]) +
  facet_grid(paste0("\u03C6: ", phi) ~ paste0("\u03C3: ", sigW))
```

##### Observation and Implementation Error

Finally, it might be worth noting that combinations of implementation and observation error do not result in additive probabilities of under- or over-escapement.

```{r echo = FALSE, warning = FALSE}
#| label: fig-sigmaFN
#| fig.height: 7
#| fig.width: 8
#| fig.cap: "Simulated percentage of under- and over-escapments for 27 possible stock-recruit parameter combinations when observation error and implemntation error covary over 3 possible value2."
#| layout-nrow: 2
#| fig-subcap: 
#|   - "Under-escapement"
#|   - "Over-escapement"
temp %>%
  filter(sigN == sigF) %>%
  filter(eg_range == "under") %>%
  ggplot(aes(x = as.character(sigN), y = pct, group = lnalpha, fill = as.character(lnalpha)))  +
  geom_bar(stat = "identity", position = "dodge") +
  labs(fill = "log(\u03B1)", y = "Percentage") +
  xlab(~ paste(sigma[F], " = ", sigma[N])) +
  facet_grid(paste0("\u03C6: ", phi) ~ paste0("\u03C3: ", sigW))

temp %>%
  filter(sigN == sigF) %>%
  filter(eg_range == "over") %>%
  ggplot(aes(x = as.character(sigN), y = pct, group = lnalpha, fill = as.character(lnalpha))) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(fill = "log(\u03B1)", y = "Percentage") +
  xlab(~ paste(sigma[F], " = ", sigma[N])) +
  facet_grid(paste0("\u03C6: ", phi) ~ paste0("\u03C3: ", sigW))
```
