---
title: "SOC_Introduction"
author: "Adam Reimer"
format:
  html:
    embed-resources: true
editor: visual
---

```{r include = FALSE}
packs <- c("tidyverse", "ggforce", "RcppRoll", "knitr")
lapply(packs, require, character.only = TRUE)

# source functions
function_files <- list.files(path="..\\functions")
lapply(function_files, function(x) source(paste0("..\\functions\\", x)))

vec_lnalpha <- seq(1, 2, length.out = 3)
vec_sigW <- seq(0.25, 0.75, length.out = 3)
vec_phi <- seq(0, .7, length.out = 3)
beta <- 0.0001
df_power <- data.frame(lnalpha = vec_lnalpha,
                       power = c(0.6, 0.7, 0.8)) #how to pick these. High for SOC sims. Should be lower for OPY sims.
input <- 
  expand.grid(lnalpha = vec_lnalpha, sigW = vec_sigW, phi = vec_phi) %>%
  mutate(beta = beta, 
         Smsy = get_Smsy(lnalpha, beta, TRUE, sigma = sigW, phi = phi),
         lb_eggers = Smsy * 0.8, #Egger bounds
         ub_eggers = Smsy * 1.6) %>%
  rowwise() %>%
  mutate(lb_pctMSY = optimise(get_bounds, #'true' OYP bounds
                              1:Smsy, 
                              lnalpha = lnalpha, 
                              beta = beta,
                              pct_MSY = 0.9,
                              correct = TRUE,
                              sigma = sigW,
                              phi = phi)$minimum,
          ub_pctMSY = optimise(get_bounds, 
                              Smsy:(Smsy*5), 
                              lnalpha = lnalpha, 
                              beta = beta,
                              pct_MSY = 0.7,
                              correct = TRUE,
                              sigma = sigW,
                              phi = phi)$minimum) %>%
  ungroup() %>%
  left_join(df_power, by = "lnalpha")

Chinook_age <- c('3' = 0.1, '4' = 0.2, '5' = 0.3, '6' = 0.38, '7' = 0.02)

input_sims = 1000
```

The stock of concern process included in the Sustainable Salmon Policy has been identifed as in need of revision. Herein I present a simulation framework to help inform and evaluate this policy revision. This document has several goals:

-   To suggest new definitions for yield and management concerns that clarify both the reason for the designation and the Department's responsibility once a designation is recommended to the Board.

-   Establish a connection between SOC designations and reductions in the productivity of a salmon stock.

-   Evaluate the appropriate criteria ADF&G should use to define *chronic inability*.

-   Establish recommendations for rebuilding a stock which has been declared a stock of concern.

-   Demonstrate the sensitivity of salmon population dynamics to the shape of the spawner-recruit relationship at low levels of population abundance.

Let's begin by considering the codified definitions within the Sustainable Salmon Policy and compare those to how the policy has be implemented over time.

## Codified SOC Definitions

::: {.callout-note appearance="simple" icon="false"}
### Concern Definitions

Yield concern means a concern arising from a chronic inability, despite the use of specific management measures, to maintain expected yields, or harvestable surpluses, above a stocks escapement needs: a yield concern is less severe than a management concern, which is less severe than a conservation concern.

Management concern means a concern arising from a chronic inability, despite the use of specific management measures, to maintain escapements for a salmon stock within the bounds of the SEG, BEG, OEG or other specified management objectives for a fishery: a management concern is not as severe as a conservation concern.

Conservation concern means a concern arising from a chronic inability, despite the use of specific management measures, to maintain escapements for a stock above a sustainable escapement threshold (SET): a conservation concern is more severe than a management concern.
:::

In practice, the definition for yield concern has proved unworkable. One issue is that it's not clear how it differs from the definition for management concern. One interpretation would be that a yield concern covers situations when we failed to have available yield over the top end of the escapement goal. While that would be a clean definition it contradicts both the theory and practice of escapement goal management in which we strive to distribute escapements throughout the escapement goal range. In contrast, while the definition of a management concern references bounds (implying over-escapement could be a management concern) we have always interpreted it to mean under-escapement. But this use collapses the definitions for yield and management concern and contradicts the implied order of severity. Over time, ADF&G has transitioned to declaring management concerns exclusively and stopped recommending yield concerns.

In contrast, the definition of a conservation concern is very clear, provided a sustainable escapement threshold has been established. In practice, the department has yet to set an SET which negates our ability to recommend a stock of conservation concern and may leave us venerable to other forms on conservation-based management.

::: {.callout-note appearance="simple" icon="false"}
### SET Definition

"sustainable escapement threshold" or "SET" means a threshold level of escapement below which the ability of the salmon stock to sustain itself is jeopardized: in practice, an SET can be estimated based on lower ranges of historical escapement levels, for which the salmon stock has consistently demonstrated the ability to sustain itself: the SET is lower than the lower bound of the BEG and lower than the lower bound of the SEG: the SET is established by the department in consultation with the board, as needed, for salmon stocks of management or conservation concern.
:::

Unfortunately this SET definition is inconsistent with one of the fundamental assumptions in salmon stock assessment. A future chapter will discuss this definition and how the department may go about setting an SET which would allow department staff, the board and the public to have better understanding of when a stock's viability is in danger.

All three concern definitions are dependent on demonstrating a chronic inability to achieve a management objective. The definition of chronic inability is vague and had led to different interpretations at the division / regional level throughout the department.

::: {.callout-note appearance="simple" icon="false"}
### Chronic Inability

Chronic Inability means the continuing or anticipated inability to meet escapement thresholds over a four to five year period, which is approximately the generation time of most salmon species.
:::

It's likely that this definition was left deliberately vague to enable latitude to include professional judgement in SOC recommendations. In practice, I suggest there remains considerable latitude to recommend, or fail to recommend, a stock as a SOC without also needing a poorly defined threshold. That said, these simulations should allow us to quantify the utility of different SOC thresholds and hopefully coalesce around one with favorable properties.

The SOC policy could also use clarification regarding the department's response to a SOC designation... i.e. the action plan.

::: {.callout-note appearance="simple" icon="false"}
## Action Plan Components

\(4\) in association with the appropriate management plan, the department and the board will, as appropriate, collaborate in the development and periodic review of an action plan for any new or expanding salmon fisheries, or stocks of concern: action plans should contain goals, measurable and implementable objectives, and provisions, including

(A) measures required to restore and protect salmon habitat, including necessary coordination with other agencies and organization:
(B) identification of salmon stock or population rebuilding goals and objectives:
(C) fishery management actions needed to achieve rebuilding goals and objectives, in proportion to each fishery's use of, and hazards posed to, a salmon stock:
(D) description of new or expanding salmon fisheries, management concern, yield concern, or conservation concern: and
(E) performance measures appropriate for monitoring and gauging the effectiveness of the action plan that are derived from the principles and criteria contained in this policy.

\(5\) each action plan will include a research plan as necessary to provide information to address concerns; research needs and priorities will be evaluated periodically, based on the effectiveness on the monitoring described in (4) of this subsection
:::

This section of the Sustainable Salmon Policy demonstrates that that the policy, the stock of concern process, and the resultant action plan were conceived with an eye towards historical threats to fisheries (i.e. habitat degradation, poor hatchery practices, and over-fishing) and is poorly positioned to respond to reduced run sizes driven by factors largely outside of our control. I'll suggest that the majority of our action plans struggle with the issue of what they should contain... generally speaking we have healthy habitat, responsible hatchery practices, and have exhausted all of our fishery management options by the time a stock of concern recommendation is being considered.

That said, the action plan section does discuss 2 concepts that are perhaps underutilized by ADF&G staff; the idea of rebuilding goals and objectives and the idea of a research plan. These simulations will explore both issues with an eye towards management and research actions that should be implemented after a SOC has been recommended.

## Proposed SOC Definitions

The definitions of yield and management concern could be reworded to make a distinction between harvest related concerns and concerns which are outside of human control.

::: {.callout-tip appearance="simple" icon="false"}
### Proposed Concern Definitions

Yield concern means a concern arising from a chronic inability, despite the use of specific management measures, to **maintain escapements for a salmon stock above the lower bound** of the SEG, BEG, OEG or other specified management objectives for a fishery: a yield concern is not as severe as a management concern.

Management concern means a concern arising from a chronic inability to **realize annual run sizes for a salmon stock above the lower bound** of the SEG, BEG, OEG or other specified management objectives for a fishery: a management concern is not as severe as a conservation concern.
:::

These definitions clearly distinguish between the two concern types, clarify the cause of each concern recommendation, and provide guidance about the responsibilities of both ADF&G and the board in the SOC process. The definition of a yield concern (S \< escapement goal lower bound and N \> escapement goal lower bound) implies that there is harvest which could be curtailed to rebuild the stock. In this instance an action plan would be highly appropriate because over harvest is responsible for the recommendation and either inseason management or the management plan need to change to correct the situation.

The definition of a management concern (N \< escapement goal lower bound) makes it clear that we have exhausted our management capability. ADF&G should be transparent about this. In this instance[^1] a research plan is likely all that is necessary and appropriate. The goal of the research plan should be explicit: to improve or maintain stock assessments so that the state possesses the information necessary to determine an SET for the stock. Using the definition of an SET this means we need to be paying for stock assessments robust enough to estimate returns from each spawning event. Therefore, SETs (and thus conservation concerns) would be determined on a case-by-case basis for individual stocks.

[^1]: Assuming there is no obvious action we can take (i.e. a habitat issue we have to tools to address).

We should test our ability to reliably identify an SET threshold but for the moment let's define an SET as $1/2$ of the lower bound of the escapement goal and use this definitions in the simulation model.

## Parameter Space for Salmon Populations

As stated previously we will use simulation to describe how salmon populations will respond to the stock of concern definitions described above. In an effort to represent a wide variety of species and populations simulations will be performed across a grid of stock-recruit parameters. We will assume salmon population dynamics are described by a Ricker curve[^2] where the parameter space is inclusive of $\textrm{log}(\alpha)$ between $1$ and $2$, process error $\sigma$ between $0.25$ and $0.75$, and autocorrelation $\phi$ between $0$ and $0.7$. We plan to simulate 27 parameter combinations using all combinations of 3 points within each parameter's range (minimum, midpoint, maximum) while setting $\beta=0.0001$ for all simulations. The Ricker curves associated with all 27 of these parameter combinations can be seen as the black lines in @fig-Ricker2.

[^2]: Until we get to the section on conservation.

### Productivity Regimes {#sec-productivity-regimes}

```{r include = FALSE}
Ricker_plots_Seq <- list()
for(i in 1:length(unique(input$lnalpha))){
  Ricker_plots_Seq[[i]] <- 
    input %>% 
    slice(rep(1:n(), each = 251)) %>% 
    mutate(S = rep(seq(0, 5 * 1 / max(beta), by = 200), times = nrow(input)),
           R = Ricker(lnalpha = get_lnalpha_p(lnalpha, beta, sigW, phi), beta = beta, S = S),
           Rmsy = Ricker(get_lnalpha_p(lnalpha, beta, sigW, phi), beta, Smsy),
           Smax = 1 / beta,
           Rmax = Ricker(lnalpha = get_lnalpha_p(lnalpha, beta, sigW, phi), beta = beta, S = Smax),
           RSeq = Ricker(lnalpha = lb_pctMSY * beta, beta = beta, S = S),
           Rmax_Seq = Ricker(lnalpha = lb_pctMSY * beta, beta = beta, S = Smax)) %>%
    ggplot(aes(x = S, y = R)) +
    geom_rect(aes(xmin = lb_pctMSY, xmax = ub_pctMSY, ymin = -Inf, ymax = Inf), fill = "grey95") +
    geom_line() +
    geom_abline(intercept = 0, slope = 1, linetype = 2) +
    geom_point(aes(x = Smax, y = Rmax)) +
    geom_segment(aes(x = Smsy, xend = Smsy, y = Smsy, yend = Rmsy), linetype = 3) +
    geom_line(aes(y = RSeq), color = "red") +
    geom_point(aes(x = Smax, y = Rmax_Seq), color = "red") +
    theme_bw() +
    facet_grid_paginate(paste0("\u03C6: ", phi) ~ paste0("ln(\u03B1): ", lnalpha) + paste0("\u03C3: ", sigW), 
                        #scales = "free_y",
                        ncol = 3, nrow = 3, page = i)
}

#pecent reduction in lnalpha
dat_pctred <- 
  input %>% 
  mutate(lnalpha_p = get_lnalpha_p(lnalpha, beta, sigW, phi),
         lnalpha_p_red = lb_pctMSY  * beta,
         pct_red = lnalpha_p_red / lnalpha_p)
mean(dat_pctred$pct_red)
median(dat_pctred$pct_red)
range(dat_pctred$pct_red)
```

Our collective best guess about why certain salmon stocks are struggling recently is that marine productivity has undergone a large decline. Since our interest is in the steady-state behavior of our SOC definitions and associated management actions it would be convenient to describe the stock-recruit relationship under both productivity regimes. I'm going to suggest that the current population dynamics for some of our struggling stocks is best described by holding the $\beta$ parameter constant and reducing $\textrm{log}(\alpha)$ such that $\textrm{log}(\alpha)^*=\frac{lb}{\beta}$ where $lb$ is the lower bound of the escapement goal range. This situation is represented by the red lines in @fig-Ricker2. For labeling and discussion purposes we will continue to describe parameter combinations based on historical productivity and distinguish low productivity simulations with the words "low productivity". Notice that under these marine productivity conditions there is little harvestable surplus and the stock replaces itself at the lower bound of the escapement goal range. Also notice that maximizing recruitment would be strikingly inefficient.

This may be a reasonable representation of current salmon population dynamics. First, under reduced productivity we expect a stable equilibrium at the lower bound of the current escapement goal which fits that pattern we are seeing, i.e with no harvest total run continues to approximate the lower bound of the escapement goal[^3]. Second, $\beta$ is generally described as a measure of density dependence. It seems reasonable to suspect this event occurs in freshwater (when modeling a single species stock recruit relationship) where density would be highest. Note that this assumption may be difficult to verify empirically because SR pairs under low productivity regimes would appear as high leverage points in a linearized Ricker despite being produced by a change in the intercept $\textrm{log}(\alpha)$ rather than the slope $\beta$ of the stock recruit relationship. Lastly, an assumption of constant $\beta$ encourages conservative management because it implies minimal harvestable surplus below the lower bound of the goal (which appears to be true) rather than tempting mangers to chase harvestable surplus at continuously decreasing estimates of $S_{MSY}$.

[^3]: Assuming a stable age at maturity and the mean return equal the lower bound of the escapement goal and actual return equal the lower bound of the escapement goal plus process error.

```{r, echo = FALSE, fig.height=8, fig.width=8}
#| label: fig-Ricker2
#| fig.height: 7
#| fig.width: 8
#| fig.cap: "Stock recruit relationships and escapement goal ranges used to simulate Alaskan salmon population dynamics under 2 productivity regimes. The solid black line represents the mean return, dotted dashed line represents MSY, the dot represents R~max~, and the shaded area represents the escapement goal range under historical ocean productivity conditions. The red line / dot show the same parameters under low marine productivity. The dashed line represents the 1:1 line."
#| layout-nrow: 3
#| fig-subcap: 
#|   - "log(\u03B1) = 1.0"
#|   - "log(\u03B1) = 1.5"
#|   - "log(\u03B1) = 2.0"
Ricker_plots_Seq[[1]] + coord_cartesian(xlim = c(0, 20000), ylim = c(0, 20000))
Ricker_plots_Seq[[2]] + coord_cartesian(xlim = c(0, 30000), ylim = c(0, 30000))
Ricker_plots_Seq[[3]] + coord_cartesian(xlim = c(0, 50000), ylim = c(0, 50000))
```

```{r, include = FALSE}
dat_pctred <- 
  input %>% 
  mutate(lnalpha_p = get_lnalpha_p(lnalpha, beta, sigW, phi),
         lnalpha_p_red = lb_pctMSY  * beta,
         lnalpha_pctred = lnalpha_p_red / lnalpha_p,
         MSY_p = get_MSY(lnalpha, beta, correct = TRUE, sigma = sigW, phi = phi),
         MSY_p_red = get_MSY(lb_pctMSY  * beta, beta, correct = FALSE),
         MSY_pctred = MSY_p_red / MSY_p) %>%
  select(lnalpha, ends_with("_pctred")) %>%
  group_by(lnalpha) %>%
  summarise(across(ends_with("_pctred"), list(median = median, min = min, max = max)))
```

Under these assumptions of low productivity realized $\textrm{log}(\alpha)$ is reduced to `r round(dat_pctred$lnalpha_pctred_median[dat_pctred$lnalpha==1], 2)`, `r round(dat_pctred$lnalpha_pctred_median[dat_pctred$lnalpha==1.5], 2)`, and `r round(dat_pctred$lnalpha_pctred_median[dat_pctred$lnalpha==2], 2)` of the historical level for $\textrm{log}(\alpha)$ of 1, 1.5 and 2, respectively, while $MSY$ is reduced to `r round(dat_pctred$MSY_pctred_median[dat_pctred$lnalpha==1], 2)`, `r round(dat_pctred$MSY_pctred_median[dat_pctred$lnalpha==1.5], 2)`, and `r round(dat_pctred$MSY_pctred_median[dat_pctred$lnalpha==2], 2)` of historical levels of $MSY$ when historical $\textrm{log}(\alpha)$ was 1, 1.5 and 2, respectively.
